# Practical Machine Learning: Personal Activity
```{r echo=FALSE}
    library(caret)
library(randomForest)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Loading and preprocessing the data
The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
train_pml <- read.csv('pml-training.csv')
test_pml <- read.csv('pml-testing.csv')
```

##Preprocessing
### Making a training set and validation set

The trainging set (train_pml) is splits it in two sets for trining the model and validate it later-  
```{r}
set.seed(6481)
indexes <- createDataPartition(train_pml$classe, p = 0.7, list = FALSE)
set_tra <- train_pml[indexes , ]
set_val <- train_pml[-indexes , ]
```
### Clean the training set
First clean columns with zero variance near. Later exclude form model columns with more of 40% of NA values and columns with character type
```{r}
zerocol <- nearZeroVar(set_tra )
set_tra <- set_tra [, -zerocol ]

cntlength <- sapply(set_tra , function(x) {
    sum(!(is.na(x) | x == ""))
})
nullcol <- names(cntlength[cntlength < 0.6 * length(set_tra$classe)])
descriptcol <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")
excludecols <- c(descriptcol, nullcol)
set_tra  <- set_tra [, !names(set_tra ) %in% excludecols]
```

##Building the model
The model is building with randomForest package.
```{r}
rfModel <- randomForest(classe ~ ., data = set_tra , importance = TRUE, ntrees = 10)
```
Check the perfomance from training and the cross validation set.
```{r}
ptmodel <- predict(rfModel, set_tra )
print(confusionMatrix(ptmodel, set_tra $classe))
```
##Model Validation

Let us now see how our model performs on the cross validation set that we held out from training.
```{r}
pvalidation <- predict(rfModel, set_val )
print(confusionMatrix(pvalidation, set_val $classe))
```
The cross validation accuracy is 0.9959, then  our model performs rather good.

##Predicting in the Test set
With the 20 cases in the test file (pml-testing.csv) We have the follow results:
```{r}
fortest <- predict(rfModel, test_pml)
fortest
```

